A Dearth of Young Stars and Planets in the Kepler Field

### 2024-05-28 paper/ms.tex
update ms

		\vspace{-0.3cm}
		\subfloat{
			\includegraphics[width=0.475\textwidth]{st_params_adopted_logg_vs_adopted_Teff.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
  \caption{{\bf The stars.}  Our analysis focuses on stars observed by
  Kepler with previously reported rotation periods (gray points).  The
  rotation periods are primarily drawn from
  \citet{Santos_2019,Santos_2021}.  About half of these rotators are
  suitable for gyrochronology (dark gray points), based on factors
  including their non-binarity and proximity to the spectroscopic main
  sequence (orange boundary; see Section~\ref{subsec:flags}).  Some
  are ``confirmed'' or ``candidate'' KOIs that meet an additional set
  of planetary quality criteria (yellow points;
  Section~\ref{subsec:plflags}).  Surface gravities and effective
  temperatures were derived photometrically by
  \citet{Berger_2020a_catalog}.  Isochrones in the lower panel are
  from MIST \citep{Choi_2016}.
	}
	\label{fig:stellarprops}
\end{figure}

@@

{\it Temperature range (bit 0)}---We require stars to have effective
temperatures $T_{\rm eff}$ between 3800 and 6200\,K in order to report
gyrochrone ages \citep{Bouma_2023}.   Stars hotter than 6200\,K spin
down very slowly, if at all.  Stars cooler than 3800\,K do spin down
over gigayear timescales
\citep{2016ApJ...821...93N,2023ApJ...954L..50E,2024arXiv240312129C},
but the currently available open cluster data have yet to clarify
exactly when the intrinsic scatter in the population decreases.

@@

{\it Near the main sequence (bit 9)}---Many stars observed by Kepler
are far from the main sequence: this can be seen in both panels of
Figure~\ref{fig:stellarprops}.  This is a challenge for rotation-based
ages because {\it i)} point sources far from the main sequence include
unresolved photometric binaries, and {\it ii}) effects other than
magnetic braking influence the rotation periods of bona fide single
stars far from the main sequence.  The added concerns of
reddening and metallicity further complicate matters.  After exploring
various options, we settled on the orange locus in the $\log
g$--$T_{\rm eff}$ plane shown in Figure~\ref{fig:stellarprops}.  While
the exact details of how this locus is constructed are arbitrary (see Appendix~\ref{app:linemethod}),
the general aim is to flag stars for which there is evidence based on
their location in the $\log g$--$T_{\rm eff}$ plane that our
gyrochronology model may not be reliable.  
%The rotation period
%distribution of sources excluded in this manner includes a relatively
%larger fraction of ``rapid rotators'' than the main sequence sources,
%likely due to a combination of astrophysical and systematic effects.
See for instance Figure~8a of \citet{2022AJ....164..137K} or
Figures~12 and 13 of \citet{2023ApJS..268....4F} for examples of how
the period distribution of this population differs from stars near the
main sequence.

>>>>

		\vspace{-0.3cm}
		\subfloat{
			\includegraphics[width=0.475\textwidth]{st_params_adopted_logg_vs_adopted_Teff.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
  \caption{{\bf The stars.}  Our analysis focuses on stars observed by
  Kepler with previously reported rotation periods (gray points).  The
  rotation periods are primarily drawn from
  \citet{Santos_2019,Santos_2021}.  About half of these rotators are
  suitable for gyrochronology (dark gray points), based on factors
  including their non-binarity and proximity to the main
  sequence (see Section~\ref{subsec:flags}).  Some
  are ``confirmed'' or ``candidate'' KOIs that meet an additional set
  of planetary quality criteria (yellow points;
  Section~\ref{subsec:plflags}).  Surface gravities and effective
  temperatures were derived photometrically by
  \citet{Berger_2020a_catalog}.  Isochrones in the lower panel are
  from MIST \citep{Choi_2016}.
	}
	\label{fig:stellarprops}
\end{figure}

@@

{\it Temperature range (bit 0)}---We require stars to have effective
temperatures $T_{\rm eff}$ between 3800 and 6200\,K in order to report
gyrochrone ages \citep{Bouma_2023}.   Stars hotter than 6200\,K spin
down very slowly, if at all.  Stars cooler than 3800\,K do spin down
over gigayear timescales
\citep{2016ApJ...821...93N,2023ApJ...954L..50E,2024arXiv240312129C},
but the currently available open cluster data have yet to clarify
when the intrinsic scatter in the population decreases.

@@

{\it Near the main sequence (bit 9)}---Many stars observed by Kepler
are far from the main sequence: this can be seen in both panels of
Figure~\ref{fig:stellarprops}.  Some of the challenges this introduces
for rotation-based ages include unresolved binaries, metallicity,
reddening, and spin-evolution effects other than magnetic braking.
After exploring various options, we settled on the orange locus in the
$\log g$--$T_{\rm eff}$ plane shown in Figure~\ref{fig:stellarprops}
as a way of flagging unresolved binaries, as well as mildly evolved
late-F and early-G stars.  While the exact details of how this locus
is constructed are arbitrary (see Appendix~\ref{app:linemethod}), the
general aim is to flag stars for which there is evidence based on
their location in the HR diagram that our gyrochronology model may not
be reliable.  While this $\log g$--$T_{\rm eff}$ cut flags about half
of Kepler stars with rotation periods as potentially not being
suitable for gyrochronology, a few outliers in $M_{\rm G}$ vs.~$G_{\rm
BP}$-$G_{\rm RP}$ remained after applying this cut.  Such stars are
likely also questionable.  We therefore also
fitted a polynomial to the KOI main sequence, and flagged stars more
than 1\,magnitude from this locus as part of the same bitflag.

%The rotation period
%distribution of sources excluded in this manner includes a relatively
%larger fraction of ``rapid rotators'' than the main sequence sources,
%likely due to a combination of astrophysical and systematic effects.
%See for instance Figure~8a of \citet{2022AJ....164..137K} or
%Figures~12 and 13 of \citet{2023ApJS..268....4F} for examples of how
%the period distribution of this population differs from stars near the
%main sequence.
